Food-web structure, stochasticity and response diversity drive temporal community stability in complex communities

### 2024-01-15 paper/manuscript.Rmd
Cleaning for sending to coauthors

---
title: Food-web structure, stochasticity and response diversity drive temporal community stability in complex communities
author: "Alain Danet, (Sonia Kéfi, Thomas F Johnson), Andrew P. Beckerman"
date: ""
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
    keep_tex: true
    spacing: double
    number_sections: false
  officedown::rdocx_document:
    base_format: "bookdown::word_document2"
    number_sections: false
always_allow_html: true
bibliography: references.bib
link-citations: true
toc: false
linestretch: 2.0
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \usepackage{hyperref}
  - \hypersetup{backref=true}
  - \usepackage{setspace}\doublespacing
---

@@

First, the stability of communities is often evaluated using temporal
variability the metric of interest (e.g. abundance or biomass), because most
natural communities face persistent perturbations so that other stability
metrics such as resilience and resistance are not always applicable
[@arnoldi_resilience_2016]. Understanding temporal community variability,
measured as the inverse of the coefficient of variation
[@tilman_biodiversity_2006; @roscher_identifying_2011;
@thibaut_understanding_2013; @zhao_biodiversity_2022], requires to scale down
the focus to populations. Community stability can be mathematically partitionned
into average population stability and asynchrony [@loreau_species_2008;
@thibaut_understanding_2013; @zhao_biodiversity_2022], asynchrony measuring how
much each species fluctuations are compensating each other at the community
level. Population stability and asynchrony depend both on the number of species
and their dominance [@loreau_species_2008; @thibaut_understanding_2013; @zhao_biodiversity_2022],
the interactions among them [@thebault_trophic_2005;
@thebault_relationship_2006; @loreau_species_2008; @danet_species_2021;
@eschenbrenner_diversity_2023] and on the way that species respond to
environmental perturbations [response diversity, @ives_stability_2000;
@elmqvist_response_2003; @ross_how_2023]. Therefore, accounting for the effects
on community structure on population stability and asynchrony is crucial to
improve our understanding of community stability in face of environmental
change.

Average population stability is decreased by external forcing, such as the
intensity and the frequency of perturbations [@loreau_species_2008]. Species richness can
decrease [@tilman_biodiversity_2006; @loreau_species_2008], increase
[@houlahan_negative_2018], or have little effects on population stability. This
effect of species richness on population stability depends on the evenness of
the community such as dominant species contribute more the average of the
community than surbordinate species. At the core of ecological theory,
population stability depends on internal forcing, here species interactions. The
topogical structure and the distribution of interaction strength have tremendous
effects on population stability, such as strong interaction strength are
generally destabilising [@mccann_weak_1998; @mccann_reevaluating_1997;
@thebault_trophic_2005]. In particular for food-webs, top-down and bottom-up
controls can increase population stability.

Asynchrony in turn can be partitionned into portefolio effects and compensatory
dynamics and is mostly determine by internal factors. Portefolio effects stems
from the statistical averaging of independent population fluctuations
[@doak_statistical_1998; @zhao_biodiversity_2022], thereby increasing
asynchrony. Compensatory dynamics in turn captures population fluctuations that
are more synchronous or asynchronous than expected for independant fluctuations,
and can stem from species interactions and species responses to perturbations
[@zhao_biodiversity_2022]. Response diversity captures how among species’
differences in their response to the environmental variation, can underpin
asynchronous dynamics if their responses are different but synchronous if their
responses are similar. Theory and empirical data emerging around response
diversity suggests it is  “core ecological mechanism underpinning many of the
statistical explanations of the diversity–stability relationships”
[@ives_stability_2000; @elmqvist_response_2003; @ross_how_2023]. In turn,
the distribution and strength of species interactions can also modulate
asynchrony through compensatory dynamics such as a strong asymetry of consumer
preference can generate asynchrony in resource fluctuations [@mccann_weak_1998;
@vasseur_environmental_2007], and a generalist consumer can in turn generate
synchrony among its prey [@raimondo_interspecific_2004]. Competition can also
have synchronising effects in strong species growth rates regime
[@loreau_species_2008]. Finally, portefolio and compensatory effects can also be
affected by the population evenness such as fluctuations of a dominant
species will be hardly compensated by a subordinate species. Improving our
understanding of community stability facing perturbations therefore requires to
articulate the effect response diversity, species interactions, species
richness and evenness on the different components of stability.

@@

Our primary objective is to derive theory that predicts how the components of
community stability, that can differentially drive patterns of stability, deliver novel
insights into the source and patterns of diversity-stability relationships in
complex communities facing environmental stochasticity. Our approach delivers
revised insight into diversity-productivity relationships in species rich
communities when both community structure per se, and species
responses can vary. In the following sections, we begin by introducing the
multi-species model and the features of the model that allow us to integrate
across variation in community structure and response diversity. We
then introduce assumptions we use to implement environmental stochasticity,
response diversity, variation in community structure and variation in the
distribution of interaction types and strengths. Next we define our in-silico
experimental design that specifically aligns with three core questions and
associated analyses: (1) What are the direct effects of food-web structure and
interaction strength on population stability, compensatory dynamics,
and statistical averaging effects (i.e. portefolio effects)? (2) What are the
total effects of food-web structure, reponse diversity, environmental
stochasticity on community stability. Finally merging the two parallel bodies of
work, we asked (3) How do response diversity and food-web structure interact to
drive the sign of the relation between community stability and
species richness?

@@

The consumers gain biomass by consumming resources (i.e. their preys) at a rate
that depends on their mass specific metabolic rate ($x_i$),  maximum cumsumption
rate ($y_i$) and on the functional response ($F_ij$) describing how the rate of
consumption of a consumer $i$ on a resource $j$ vary with the biomass of this resource.
Consumers then loses biomass by being consumed, $e_ij$ being the assimilisation
efficiency of the consumer $j$ on the resource $i$. Consumers also lose biomass
over time through metabolic losses ($-x_iB_i$) and by natural mortality due to
environment ($-d_iB_i$). The primary producers gain biomass over time with a
growth rate ($r_i$, eq. \@ref(eq:prod)) and a functional response ($G_i$)
describing how the growth of the producers vary with their biomass. They lose
biomass by being consumed by consumers and by natural mortality.

@@

The growth function of the primary producers (eq. \@ref(eq:growth)) has the shape of a logistic growth
where the growth rate is maximal and null when
$\sum_{j \in \{\mathrm{prod}\}} \alpha_{ij} B_j$ are respectively close to 0 and 1. $\alpha_{ij}$ is the
per capita effect of the producer $j$ on the producer $i$, $K_i$ being the
carrying capacity for the producer $j$. The functional response of a consumer
feeding on a resource (eq. \@ref(eq:holling)) depends on the relative preference
of the consumer on the resource ($\omega_ij$) is limited by an half-saturation
rate ($B_0$), intraspecific interference coefficient ($c_i$), and on the
availability of its other resources
($\sum_{k \in \{\mathrm{res}\}_i} \omega_{ik} B_k^h$). Finally, the $h$ exponent
controls the shape of the functional response from a saturating function
(Holling type II) to a sigmoid (Holling type III).

@@

We introduced environmental stochasticity by
considering stochastic variation of mortality rates. Doing so allows to consider
stochasticity without modifying species interactions themselves
[@vasseur_environmental_2007]. The stochastic part of the mortality rates
($\epsilon_{i, t}$) follow a normal distribution 0 centered
($\epsilon_{i, t} \sim N(0,\sigma^2_e)$) with a variance ($\sigma^2_e$). The mortality rate each
species at the time $t$ equals to $d_{i,t} = d_i e^{\epsilon_{i, t}}$. We did so
to ensure to not have negative mortality rates [@vasseur_environmental_2007]. We
simulated $\epsilon$ using an Ornstein-Uhlenbeck process, which is a modified
browian motion where the stochastic values tend to come back to the central
value (i.e. $\epsilon = 0$) such as $\epsilon_t$ varies according to this
stochastic differential equation: $d\epsilon_t = (0 - \epsilon_t)d_t + \sigma_e
dW_t$, where $dW_t$ is a Brownian motion. The strength of environmental
stochasticity was then controlled by $\sigma_e$.

@@

where $a_{i,t}$ the stochastic value generated by the Ornstein-Uhlenbeck
process at time $t$, $\rho_{ii} = 1.0$ while all $\rho_{ij, i \neq j}$ were equal and control
the level of response diversity.

@@

```{r}
tar_read(tab_sim_param) %>%
  kable(
    booktabs = TRUE,
    caption = "Parameter values used in the simulation experiment. Numbers
    between squared brackets display all the values. Two numbers between
    brackets are the minumum and maximal values, N being the number of unique
    values. PPMR: Predator Prey Mass Ratio. We defined response diversity as \\protect$1-\\rho$
    (ref:eq-rho).",
    label = "param") %>%
  kable_styling(latex_options = c("striped", "hold_position"))
```

@@

To vary interaction strength and mortality rates (i.e. allometrically
scaled mortality rate $d$) in the community, we varied the Predator-Prey
Bodysize Ratio (PPBR) from $`r pZ["min"]`$ to $`r pZ["max"]`$ (Fig. \@ref(fig:met)b). We further varied
the functional response of consumer from type II to type III (resp.
$`r ph["min"]`$ to $`r ph["max"]`$, eq. \@ref(eq:holling)) as it was shown to
have a tremendous role on stability [@brose_allometric_2006]. We set a global
carrying capacity of $K' = 10$ to ensure that consumers were not limited by the
biomass input from primary producers. We standardised the carrying capacity by
the number of primary producers and by the interspecific competition among
producers to ensure that the effects of species richness on temporal stability
are not driven by the increase in the number of primary producers, thereby
trivially increase biomass input for consumers. We then standardised $K'$ such as
$K_i = \frac{1 + \alpha_{ij} (S - 1)}{S}K'$ [@ives_stability_2000].

@@

We measured food-web structure and dynamics that have been shown to be linked
with stability. We collected total biomass, species richness, connectance,
average trophic level weighted by biomass, average omnivory and average per
capita interaction strength. Connectance was computed as $C = L / S(S-1)$, $L$
being the number of trophic links and $S$ being the number of species in the
community. Species trophic level was computed recursevily from the bottom of the
food-web, such as the trophic level of a consumer was equal to the average
trophic level of its resources added to 1, the trophic level of primary
producers being set to 1 [@christensen_ecopath_1992]. Average trophic level was then equal to:
$1/\hat{B}_i\sum_{i}\mathrm{TL}_i$, $\mathrm{TL}_i$ being the trophic level and
$\hat{B}_i$ the average biomass of the species $i$. The degree of omnivory of a
consumer was computed  such as the sum of square of its
resource trophic levels weighted by the relative preference of the consumer for
this resource: $\sum_i (\mathrm{TL}_i - \hat{\mathrm{TL}})^2 \omega_i/ \sum \omega_i$ [@christensen_ecopath_1992].
Interaction strength was computed was quantified as the biomass fluxes going
from a resource to a consumer, such as: $I = x_i y_i B_k F_{ij}$ (eq.
\@ref(eq:cons)), and averaged over time. Per capita interaction strength was
obtained by dividing average interaction strength by the average biomass of the
consumer. Finally, the average interaction strength of a community was computed
as the average per capita interaction strength at the exclusion of null
interaction strengths (i.e. non present trophic interactions).

@@

As SAE is further dependant on the evenness of species variance, we
partitioned $SAE$ into $SAE_{even}$ and evenness $EVN$, such as  $EVN = SAE / SAE_{even}$.
$SAE_{even}$ corresponds to the $SAE$ when species quantities are perfectly
even such as from eq. \@ref(eq:sae2), we can obtain $SAE_{even}$ from which $EVN$ can be computed:

@@

with $\mathbf{B}$ and $\mathbf{E}$ being the matrices time x species for species biomass
and stochastic mortality rates respectively. Then compensatory dynamics raising
from species interactions were computed as: $CPE_{int} = CPE / CPE_{env}$.

@@

Prior to test the structural equation model, we logged all the stability
components to transform their relation from multiplicative to additive (eq XX).
We ensure that all the linear models composing the structural equation model
presented low multicollinearity (Variance Inflation Factor < 5, Table SXX). The
structural equation model using the R package `PiecewiseSEM`
[@lefcheck_piecewisesem_2016]. We then bootstrapped (N = 100) the SEM to obtain
confidence interval on the estimated effects and computed the sum of the direct
and indirect effects using `semEff` R package [@murphy_semeff_2021]. In the
main text, we reported standardized coefficients which are obtained by scaling
the coefficients by the stardard deviation of the response and predictor
variable (add formula). Finally, because we analyzed simulation, we reported
only the direct standardized coefficients that had an absolute value above $0.1$.

In a second analysis, we aimed to test how stability-species richness
relationships were modulated by food-web structure and the two stressors:
environmental stochasticity and response diversity. We modelled temporal
stability of biomass according to food-web structure: species richness, weighted
average trophic level, connectance, hill exponent, predator-prey size ratio, and
average interaction strength, as well as environmental stochasticity and species
response diversity. We further included the two way interactions between species
richness and food-web structure, plus between species richness and stressors.
Finally we added the three way interaction between species richness, food-web
structure and stressors. The VIF were inferior to 5, indicating low
multicollinearity. We then used the coefficients of the model to predict the shape
of the relationship between community stability and species richness according
to a gradient of response diversity and of food-web structure.

@@

We found that food-web structure had significant effect on temporal population
stability such as average trophic level and type III functional response (i.e.
hill exponent = 3) increase population stability (resp. $r_\delta$= `r rpt` and
`r rph`, $r_\delta$ being the standardized coefficient, Fig. \@ref(fig:sem)a).
In turn, the variance of environmental stochasticity had a strong negative effect
on population stability ($r_\delta$= `r rpe`).

@@

Deriving the sum of direct and indirect effects from the structural equation
model, we show that both environmental stochasticity and response diversity among
species had by far the most important effect on community stability (resp.
$r_\delta$= `r rtsenv` and `r rtsdiv`, Fig. \@ref(fig:sem)b). Species richness
had a low overall positive total effect on community stability (`r rtsr`)
because its strong total positive effect on asynchrony (`r rtar`) was downplayed
by its total negative effect on population stability (`r rtpr`) and that the
weight of population stability on community stability is twice higher than the
one of asynchrony. Average trophic level, type III functional response and
predator prey bodysize ratio had positive effects on community stability in
total (resp. `r rtst`, `r rtsh`, `r rtsz`, Fig. \@ref(fig:sem)b), driven by
their positive effects on population stability that were weakly counterbalanced
by their negative effects on asynchrony. In turn, connectance and average
interaction strength had low total negative effects on community stability (resp.
`r rtsc` and `r rtsi`, Fig. \@ref(fig:sem)b). While connectance had negative
effects both on population stability and asynchrony, average interaction
strength had a low total positive effect on population stability counterbalanced
by a stronger negative effect on asynchrony.

@@

Our results show that community stability in complex food-webs is mainly driven
by population stability rather than synchrony, confirming previous results in
empirical fish communities [@danet_species_2021], which contrasts with empirical
data in birds and bats assemblages that reported a equivalent influence of
population stability and asynchrony on community stability
[@olivier_urbanization_2020]. Although the number of empirical and theoretical
studies is still limited to compare the respective contribution of asynchrony and
population stability to community stability in competitive assemblages and
food-webs, we suggest two non-exclusive and complementary explanations that
might drive a lower contribution of asynchrony in food-webs compared to
competitive assemblages. First, the range of bodymasses is much higher in
food-webs, which typically include species that differ in bodymass by several order of
magnitude [@brose_predator_2019], which in turn creates highly uneven biomass
distribution, thereby creating uneven contributions of species to asynchrony
[@thibaut_understanding_2013] and limits the raise of high portfolio effects and
compensatory dynamics in species rich communities [@zhao_biodiversity_2022].
Second, the trophic interactions characterising food-webs create strong
interdependence among species, which in turn limits the raise of independent
fluctuations (i.e. portfolio effects, but rise compensatory dynamics). The question of
which mechanisms and community features drive differences in the
contribution of asynchrony and population stability in food-webs and competitive
assemblages is still a open question that might be crucial to accurately
address diagnostics and conservation issues. On top of food-web structure,
we found that environmental stochasticity and environmental correlation among
species were the main drivers of community stability, suggesting that stressors
are the main drivers of temporal stability. This result is in accordance with previous
studies showing strong effects on environmental stochasticity
[@vasseur_environmental_2007] and the loss of response diversity decrease
community stability in tri-trophic food-webs [@ives_stability_2000;
@thebault_trophic_2005].

We further found that average trophic level increased average population
stability, confirming previous theoreical [@brose_allometric_2006;
@shanafelt_stability_2018; @eschenbrenner_diversity_2023] and
[@danet_species_2021] empirical results. Higher average trophic level can
stabilise populations because higher predators are themselves more stable due
to lower mortality rates with in turn makes them less sensible stochastic
perturbation and because they control the amplitude of the fluctuations of their
preys [@shanafelt_stability_2018, we can test that?]. In turn, we found that
average trophic level decreases strongly asynchrony through a negative effect on
statistical averaging effect. Further analysis showed that average trophic level
limited statistical averaging effect by decreasing the evenness of biomass
variability. It is linked to the model assumption that species of higher trophic
level have a lower overall mortality rates, similarly to previous studies
[@vasseur_environmental_2007], thereby increasing average population stability and simultaneously
synchronising populations. The fact that Predator Size Mass Ratio displays the same
effect, thereby building food-webs with bigger species having lower mortality
rates, further lends toward this hypothesis. Furthermore, average omnivory was
strongly correlated with average trophic level, suggesting that average omnivory
also stabilise population stability, in line with previous theoretical results
in small modules [@mccann_reevaluating_1997]. Our results further suggest that
higher average trophic level results in lower compensatory dynamics raising from
species interactions. It might be explained by the fact that average trophic
level is strongly associated with average omnivory, which might tamper
synchronized predator-prey oscillations that can raise compensatory dynamics
[@mccann_reevaluating_1997].

>>>>

---
title: Food-web structure, stochasticity and response diversity drive temporal community stability in complex communities
author: "Alain Danet, (Sonia Kéfi, Thomas F Johnson), Andrew P. Beckerman"
date: ""
output:
  officedown::rdocx_document:
    number_sections: false
  bookdown::pdf_document2:
    latex_engine: xelatex
    keep_tex: true
    spacing: double
    number_sections: false
always_allow_html: true
bibliography: references.bib
link-citations: true
csl: ecology-letters.csl
toc: false
linestretch: 2.0
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \usepackage{hyperref}
  - \hypersetup{backref=true}
  - \usepackage{setspace}\doublespacing
---

@@

First, the stability of communities is often evaluated using temporal
variability the metric of interest (e.g. abundance or biomass), because most
natural communities face persistent perturbations so that other stability
metrics such as resilience and resistance are not always applicable
[@arnoldi_resilience_2016]. Understanding temporal community variability,
measured as the inverse of the coefficient of variation
[@tilman_biodiversity_2006; @roscher_identifying_2011;
@thibaut_understanding_2013; @zhao_biodiversity_2022], requires scaling down
the focus to populations. Community stability can be mathematically partitioned
into average population stability and asynchrony [@loreau_species_2008;
@thibaut_understanding_2013; @zhao_biodiversity_2022], asynchrony measuring how
much each species fluctuations are compensating each other at the community
level. Population stability and asynchrony depend both on the number of species
and their dominance [@loreau_species_2008; @thibaut_understanding_2013;
@zhao_biodiversity_2022], the interactions among them [@thebault_trophic_2005;
@thebault_relationship_2006; @loreau_species_2008; @danet_species_2021;
@eschenbrenner_diversity_2023] and on the way that species respond to
environmental perturbations [response diversity, @ives_stability_2000;
@elmqvist_response_2003; @ross_how_2023]. Therefore, accounting for the effects
on community structure on population stability and asynchrony is crucial to
improve our understanding of community stability in face of environmental
change.

Average population stability is decreased by external forcing, such as the
intensity and the frequency of perturbations [@loreau_species_2008]. Species richness can
decrease [@tilman_biodiversity_2006; @loreau_species_2008], increase
[@houlahan_negative_2018], or have little effects on population stability. This
effect of species richness on population stability depends on the evenness of
the community such as dominant species contribute more the average of the
community than subordinate species. At the core of ecological theory,
population stability depends on internal forcing, here species interactions. The
topological structure and the distribution of interaction strength have tremendous
effects on population stability, such as strong interaction strength are
generally destabilising [@mccann_weak_1998; @mccann_reevaluating_1997;
@thebault_trophic_2005]. In particular for food-webs, top-down and bottom-up
controls can increase population stability.

Asynchrony in turn can be partitioned into portfolio effects and compensatory
dynamics and is mostly determined by internal factors. Portfolio effects stems
from the statistical averaging of independent population fluctuations
[@doak_statistical_1998; @zhao_biodiversity_2022], thereby increasing
asynchrony. Compensatory dynamics in turn captures population fluctuations that
are more synchronous or asynchronous than expected for independent fluctuations,
and can stem from species interactions and species responses to perturbations
[@zhao_biodiversity_2022]. Response diversity captures how among species’
differences in their response to the environmental variation, can underpin
asynchronous dynamics if their responses are different but synchronous if their
responses are similar. Theory and empirical data emerging around response
diversity suggests it is  “core ecological mechanism underpinning many of the
statistical explanations of the diversity–stability relationships”
[@ives_stability_2000; @elmqvist_response_2003; @ross_how_2023]. In turn,
the distribution and strength of species interactions can also modulate
asynchrony through compensatory dynamics such as a strong asymmetry of consumer
preference can generate asynchrony in resource fluctuations [@mccann_weak_1998;
@vasseur_environmental_2007], and a generalist consumer can in turn generate
synchrony among its prey [@raimondo_interspecific_2004]. Competition can also
have synchronising effects in strong species growth rates regime
[@loreau_species_2008]. Finally, portfolio and compensatory effects can also be
affected by the population evenness such as fluctuations of a dominant
species will hardly be compensated by a subordinate species. Improving our
understanding of community stability facing perturbations therefore requires to
articulate the effect response diversity, species interactions, species
richness and evenness on the different components of stability.

@@

Our primary objective is to derive theory that predicts how the components of
community stability, that can differentially drive patterns of stability, deliver novel
insights into the source and patterns of diversity-stability relationships in
complex communities facing environmental stochasticity. Our approach delivers
revised insight into diversity-productivity relationships in species rich
communities when both community structure per se, and species
responses can vary. In the following sections, we begin by introducing the
multi-species model and the features of the model that allow us to integrate
across variation in community structure and response diversity. We
then introduce assumptions we used to implement environmental stochasticity,
response diversity, variation in community structure and variation in the
distribution of interaction types and strengths. Next we define our in-silico
experimental design that specifically aligns with three core questions and
associated analyses: (1) What are the direct effects of food-web structure and
interaction strength on population stability, compensatory dynamics,
and statistical averaging effects (i.e. portfolio effects)? (2) What are the
total effects of food-web structure, response diversity, environmental
stochasticity on community stability. Finally, merging the two parallel bodies of
work, we asked (3) How do response diversity and food-web structure interact to
drive the sign of the relation between community stability and
species richness?

@@

The consumers gain biomass by consuming resources (i.e. their preys) at a rate
that depends on their mass specific metabolic rate ($x_i$),  maximum consumption
rate ($y_i$) and on the functional response ($F_ij$) describing how the rate of
consumption of a consumer $i$ on a resource $j$ vary with the biomass of this resource.
Consumers then loses biomass by being consumed, $e_ij$ being the assimilation
efficiency of the consumer $j$ on the resource $i$. Consumers also lose biomass
over time through metabolic losses ($-x_iB_i$) and by natural mortality due to
environment ($-d_iB_i$). The primary producers gain biomass over time with a
growth rate ($r_i$, eq. \@ref(eq:prod)) and a functional response ($G_i$)
describing how the growth of the producers vary with their biomass. They lose
biomass by being consumed by consumers and by natural mortality.

@@

The growth function of the primary producers (eq. \@ref(eq:growth)) has the shape of a logistic growth
where the growth rate is maximal and null when
$\sum_{j \in \{\mathrm{prod}\}} \alpha_{ij} B_j$ are respectively close to 0 and 1. $\alpha_{ij}$ is the
per capita effect of the producer $j$ on the producer $i$, $K_i$ being the
carrying capacity for the producer $j$. The functional response of a consumer
feeding on a resource (eq. \@ref(eq:holling)) depends on the relative preference
of the consumer on the resource ($\omega_ij$) is limited by a half-saturation
rate ($B_0$), intraspecific interference coefficient ($c_i$), and on the
availability of its other resources
($\sum_{k \in \{\mathrm{res}\}_i} \omega_{ik} B_k^h$). Finally, the $h$ exponent
controls the shape of the functional response from a saturating function
($h = 2$, Holling type II) to a sigmoid ($h = 3$, Holling type III).

@@

We introduced environmental stochasticity by
considering stochastic variation of mortality rates. Doing so allows to consider
stochasticity without modifying species interactions themselves
[@vasseur_environmental_2007]. The stochastic part of the mortality rates
($\epsilon_{i, t}$) follow a normal distribution 0 centered
($\epsilon_{i, t} \sim N(0,\sigma^2_e)$) with a variance ($\sigma^2_e$). The mortality rates each
species at the time $t$ equals to $d_{i,t} = d_i e^{\epsilon_{i, t}}$. We did so
to ensure to not have negative mortality rates [@vasseur_environmental_2007]. We
simulated $\epsilon$ using an Ornstein-Uhlenbeck process, which is a modified
Brownian motion where the stochastic values tend to come back to the central
value (i.e. $\epsilon = 0$) such as $\epsilon_t$ varies according to this
stochastic differential equation: $d\epsilon_t = (0 - \epsilon_t)d_t + \sigma_e
dW_t$, where $dW_t$ is a Brownian motion. The strength of environmental
stochasticity was then controlled by $\sigma_e$.

@@

Where $a_{i,t}$ the stochastic value generated by the Ornstein-Uhlenbeck process
at time $t$, $\rho_{ii} = 1.0$ while all $\rho_{ij, i \neq j}$ were equal and
control the level of response diversity.

@@

```{r}
tar_read(tab_sim_param) %>%
  kable(
    booktabs = TRUE,
    caption = "Parameter values used in the simulation experiment. Numbers
    between squared brackets display all the values. Two numbers between
    brackets are the minimum and maximal values, N being the number of unique
    values. PPMR: Predator Prey Mass Ratio. We defined response diversity as \\protect$1-\\rho$
    (ref:eq-rho).",
    label = "param") %>%
  kable_styling(latex_options = c("striped", "hold_position"))
```

@@

To vary interaction strength and mortality rates (i.e. allometrically
scaled mortality rate $d$) in the community, we varied the Predator-Prey
Mass Ratio (PPMR) from $`r pZ["min"]`$ to $`r pZ["max"]`$ (Fig. \@ref(fig:met)b). We further varied
the functional response of consumer from type II to type III (i.e. from
$h = `r ph["min"]`$ to $h = `r ph["max"]`$, eq. \@ref(eq:holling)) as it was shown to
have a tremendous role on stability [@brose_allometric_2006]. We set a global
carrying capacity of $K' = 10$ to ensure that consumers were not limited by the
biomass input from primary producers. We standardised the carrying capacity by
the number of primary producers and by the interspecific competition among
producers to ensure that the effects of species richness on temporal stability
are not driven by the increase in the number of primary producers, thereby
trivially increase biomass input for consumers. We then standardised $K'$ such as
$K_i = \frac{1 + \alpha_{ij} (S - 1)}{S}K'$ [@ives_stability_2000].

@@

We measured food-web structure and dynamics that have been shown to be linked
with stability. We collected total biomass, species richness, connectance,
average trophic level weighted by biomass, average omnivory and average per
capita interaction strength. Connectance was computed as $C = L / S(S-1)$, $L$
being the number of trophic links and $S$ being the number of species in the
community. Species trophic level was computed recursively from the bottom of the
food-web, such as the trophic level of a consumer was equal to the average
trophic level of its resources added to 1, the trophic level of primary
producers being set to 1 [@christensen_ecopath_1992]. Average trophic level was then equal to:
$1/\hat{B}_i\sum_{i}\mathrm{TL}_i$, $\mathrm{TL}_i$ being the trophic level and
$\hat{B}_i$ the average biomass of the species $i$. The degree of omnivory of a
consumer was computed  such as the sum of squares of its
resource trophic levels weighted by the relative preference of the consumer for
each resource: $\sum_i (\mathrm{TL}_i - \hat{\mathrm{TL}})^2 \omega_i/ \sum \omega_i$ [@christensen_ecopath_1992].
Interaction strength was computed was quantified as the biomass fluxes going
from a resource to a consumer, such as: $I = x_i y_i B_k F_{ij}$ (eq.
\@ref(eq:cons)), and averaged over time. Per capita interaction strength was
obtained by dividing average interaction strength by the average biomass of the
consumer. Our way of measuring per interaction strength is very similar to
previous metrics used in previous studies, at the difference here that we
average fluxes over time and not the maximal interaction strength
[@mccann_weak_1998]. Finally, the average interaction strength of a community
was computed as the average per capita interaction strength at the exclusion of
null interaction strengths (i.e. excluding absence of trophic interactions).

@@

As SAE is further dependent on the evenness of species variance, we
partitioned $SAE$ into $SAE_{even}$ and evenness $EVN$, such as  $EVN = SAE / SAE_{even}$.
$SAE_{even}$ corresponds to the $SAE$ when species quantities are perfectly
even such as from eq. \@ref(eq:sae2), we can obtain $SAE_{even}$ from which $EVN$ can be computed:

@@

With $\mathbf{B}$ and $\mathbf{E}$ being the matrices time x species for species biomass
and stochastic mortality rates respectively. Then compensatory dynamics raising
from species interactions were computed as: $CPE_{int} = CPE / CPE_{env}$.

@@

Prior to test the structural equation model, we logged all the stability
components to transform their relation from multiplicative to additive (eq.
\@ref(eq:scom2)).
We ensure that all the linear models composing the structural equation model
presented low to moderate multicollinearity (Variance Inflation Factor < 10, Table SXX). The
structural equation model using the R package `PiecewiseSEM`
[@lefcheck_piecewisesem_2016]. We then bootstrapped (N = 100) the SEM to obtain
confidence interval on the estimated effects and computed the sum of the direct
and indirect effects using `semEff` R package [@murphy_semeff_2021]. In the
main text, we reported standardised coefficients which are obtained by scaling
the coefficients by the standard deviation of the response and predictor
variable (add formula). Finally, because we analysed simulation, we reported
only the direct standardised coefficients that had an absolute value above $0.1$.

In a second analysis, we aimed to test how stability-species richness
relationships were modulated by food-web structure and the two stressors:
environmental stochasticity and response diversity. We then modelled temporal
stability of biomass according to food-web structure: species richness, weighted
average trophic level, connectance, hill exponent, predator-prey mass ratio, and
average interaction strength, as well as environmental stochasticity and species
response diversity. We further included the two-way interactions between species
richness and food-web structure, plus between species richness and stressors.
Finally, we added the three-way interaction between species richness, food-web
structure and stressors. The VIF were inferior to 5, indicating low
multicollinearity. We then used the linear equation of the model and associated
slope coefficients to predict the shape of the relationship between community
stability and species richness according to a gradient of response diversity and
of food-web structure. To do so, we summed all the coefficients involving species
richness in the model, i.e. the one, two, and three-way terms. The values of the
control variables were set to the values indicated in the Fig. \@ref(fig:pred).

@@

We found that food-web structure had significant effect on temporal population
stability such as average trophic level and type III functional response (i.e.
hill exponent = 3) increase population stability (resp. $r_\delta$= `r rpt` and
`r rph`, $r_\delta$ being the standardised coefficient, Fig. \@ref(fig:sem)a).
In turn, the variance of environmental stochasticity had a strong negative effect
on population stability ($r_\delta$= `r rpe`).

@@

Deriving the sum of direct and indirect effects from the structural equation
model, we show that both environmental stochasticity and response diversity among
species had by far the most important effect on community stability (resp.
$r_\delta$= `r rtsenv` and `r rtsdiv`, Fig. \@ref(fig:sem)b). Species richness
had a low overall positive total effect on community stability (`r rtsr`)
because its strong total positive effect on asynchrony (`r rtar`) was downplayed
by its total negative effect on population stability (`r rtpr`) and that the
weight of population stability on community stability is twice higher than the
one of asynchrony. Average trophic level, type III functional response and
predator prey mass ratio had positive effects on community stability in
total (resp. `r rtst`, `r rtsh`, `r rtsz`, Fig. \@ref(fig:sem)b), driven by
their positive effects on population stability that were weakly counterbalanced
by their negative effects on asynchrony. In turn, connectance and average
interaction strength had low total negative effects on community stability (resp.
`r rtsc` and `r rtsi`, Fig. \@ref(fig:sem)b). While connectance had negative
effects both on population stability and asynchrony, average interaction
strength had a low total positive effect on population stability counterbalanced
by a stronger negative effect on asynchrony.

@@

Our results show that community stability in complex food-webs is mainly driven
by population stability rather than synchrony, confirming previous results in
empirical fish communities [@danet_species_2021], which contrasts with empirical
data in birds and bats assemblages that reported an equivalent influence of
population stability and asynchrony on community stability
[@olivier_urbanization_2020]. Although the number of empirical and theoretical
studies is still limited to compare the respective contribution of asynchrony and
population stability to community stability in competitive assemblages and
food-webs, we suggest two non-exclusive and complementary explanations that
might drive a lower contribution of asynchrony in food-webs compared to
competitive assemblages. First, the range of bodymasses is much higher in
food-webs, which typically include species that differ in bodymass by several orders of
magnitude [@brose_predator_2019], which in turn creates highly uneven biomass
distribution, thereby creating uneven contributions of species to asynchrony
[@thibaut_understanding_2013] and limits the raise of high portfolio effects and
compensatory dynamics in species rich communities [@zhao_biodiversity_2022].
Second, the trophic interactions characterising food-webs create strong
interdependence among species, which in turn limits the raise of independent
fluctuations (i.e. portfolio effects, but rise compensatory dynamics). The question of
which mechanisms and community features drive differences in the
contribution of asynchrony and population stability in food-webs and competitive
assemblages is still an open question that might be crucial to accurately
address diagnostics and conservation issues. On top of food-web structure,
we found that environmental stochasticity and environmental correlation among
species were the main drivers of community stability, suggesting that stressors
are the main drivers of temporal stability. This result is in accordance with previous
studies showing strong effects on environmental stochasticity
[@vasseur_environmental_2007] and the loss of response diversity decrease
community stability in tri-trophic food-webs [@ives_stability_2000;
@thebault_trophic_2005].

We further found that average trophic level increased average population
stability, confirming previous theoretical [@brose_allometric_2006;
@shanafelt_stability_2018; @eschenbrenner_diversity_2023] and
[@danet_species_2021] empirical results. Higher average trophic level can
stabilise populations because higher predators are themselves more stable due
to lower mortality rates with in turn makes them less sensible stochastic
perturbation and because they control the amplitude of the fluctuations of their
preys [@shanafelt_stability_2018, we can test that?]. In turn, we found that
average trophic level decreases strongly asynchrony through a negative effect on
statistical averaging effect. Further analysis showed that average trophic level
limited statistical averaging effect by decreasing the evenness of biomass
variability. It is linked to the model assumption that species of higher trophic
level have a lower overall mortality rates, similarly to previous studies
[@vasseur_environmental_2007], thereby increasing average population stability and simultaneously
synchronising populations. The fact that Predator Size Mass Ratio displays the same
effect, thereby building food-webs with bigger species having lower mortality
rates, further lends toward this hypothesis. Furthermore, average omnivory was
strongly correlated with average trophic level, suggesting that average omnivory
also stabilise population stability, in line with previous theoretical results
in small modules [@mccann_reevaluating_1997]. Our results further suggest that
higher average trophic level results in lower compensatory dynamics raising from
species interactions. It might be explained by the fact that average trophic
level is strongly associated with average omnivory, which might tamper
synchronized predator-prey oscillations that can raise compensatory dynamics
[@mccann_reevaluating_1997].
